---
title: "16p HiC - Loop Reproducibility Analysis"
author: "Siddharth Reed"
date: "`r Sys.Date()`"
---
<style type="text/css">
.main-container {
  max-width: 1800px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Dependencies + Load data

```{r knitr}
library(knitr)
knitr::opts_chunk$set(
    results=FALSE,
    message=FALSE,
    warning=FALSE,
    dev=c('png', 'pdf', 'svg'),
    dpi=300
)
```
Set up file locations + load dependences
```{r dependencies}
library(here)
here::i_am('notebooks/Loop.Reproducibility.Rmd')
BASE_DIR <- here()
source(file.path(BASE_DIR, 'scripts/locations.R'))
source(file.path(SCRIPT_DIR, 'utils.data.R'))
source(file.path(BASE_DIR, 'scripts/constants.R'))
source(file.path(SCRIPT_DIR, 'utils.plot.R'))
source(file.path(SCRIPT_DIR, 'utils.annotations.R'))
source(file.path(SCRIPT_DIR, 'loops/utils.loops.R'))
library(tidyverse)
library(magrittr)
```
Generate IDR2D results on loops for all comparisons of sample groups
```{r load_idr2d}
# library(furrr)
# plan(multisession, workers=length(availableWorkers()))
idr2d.results.df <- 
    check_cached_results(
        results_file=ALL_LOOPS_IDR2D_RESULTS_FILE,
        # force_redo=TRUE,
        results_fnc=load_all_IDR2D_results
    ) %>% 
    post_process_IDR2D_results() %>% 
    standardize_data_cols()
```
Count loops by categoy
```{r count_reproducible_loops}
# idr2d.results.df
n.idr2d.results.df <- 
    idr2d.results.df %>% 
    mutate(`metric+resolve`=glue('{metric}+{resolve.method}')) %>% 
    # order is important for calc_pct() below
    count(
        # type, weight, kernel, 
        resolution,
        resolve.method, metric, `metric+resolve`,
        max.gap.bins, 
        comparison, 
        chr,
        is.loop.shared
    )
n.idr2d.results.df
```

# Compare IDR2D hyper-params {.tabset}

## Genome-Wide {.tabset}

```{r idr2d_hyperparams_barplot_gw, results='asis', fig.dim=c(10,8)}
n.idr2d.results.df %>%
    select(-c(chr)) %>% 
    calc_pct() %>%
    make_nested_plot_tabs(
        plot.fnc=plot_barplot,
        max.header.lvl=3,
        # group.cols=c('resolution', 'comparison'),
        group.cols=c('resolution'),
        x.var='max.gap.bins',
        y.var='pct',
        y.scale.mode='pct',
        fill.var='is.loop.shared',
        label.var='pct.label',
        position='stack',
        facet.row='comparison',
        facet.col='metric+resolve',
        # scales='fixed',
        # legend.position='right',
        label.size=2,
        strip.text=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()
    )
```

## Per Chr {.tabset}

```{r idr2d_hyperparams_barplot_chr, results='asis', fig.dim=c(14,10)}
n.idr2d.results.df %>%
    calc_pct() %>%
    make_nested_plot_tabs(
        plot.fnc=plot_barplot,
        max.header.lvl=3,
        group.cols=c('resolution', 'comparison'),
        x.var='max.gap.bins',
        y.scale.mode='pct',
        y.var='pct',
        fill.var='is.loop.shared',
        label.var='pct.label',
        label.size=1,
        facet.col='chr',
        facet.row='metric+resolve',
        position='stack',
        # scales='fixed',
        # legend.position='right',
        strip.text=element_text(size=10),
        axis.text.x=element_text(angle=65, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()
    )
```

# Save Filtered IDR2D Results

Save results for specified hyper-params
```{r save_filtered_results}
idr2d.results.df <-
    idr2d.results.df %>% 
    filter_loop_IDR2D_results() %T>% 
    write_tsv(FILTERED_LOOPS_FILTERED_IDR2D_RESULTS_FILE)
n.idr2d.results.df <- 
    idr2d.results.df %>% 
    count(
        resolution,
        comparison, 
        chr,
        is.loop.shared
    )
```

# Loop Reproducibilty with chosen hyper-params {.tabset}

## Genome-Wide {.tabset}

```{r idr2d_hyperparams_barplot_gw_final, results='asis', fig.dim=c(10,6)}
n.idr2d.results.df %>%
    select(-c(chr)) %>% 
    calc_pct() %>%
    make_nested_plot_tabs(
        plot.fnc=plot_barplot,
        max.header.lvl=3,
        group.cols=c('resolution'),
        x.var='comparison',
        y.scale.mode='pct',
        y.var='pct',
        fill.var='is.loop.shared',
        label.var='pct.label',
        position='stack',
        # scales='fixed',
        # legend.position='right',
        strip.text=element_text(size=10),
        axis.text.x=element_text(angle=45, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()
    )
```

## Per Chr {.tabset}

```{r idr2d_hyperparams_barplot_chr_final, results='asis', fig.dim=c(12,10)}
n.idr2d.results.df %>%
    calc_pct() %>%
    make_nested_plot_tabs(
        plot.fnc=plot_barplot,
        max.header.lvl=3,
        group.cols=c('resolution'),
        x.var='chr',
        y.scale.mode='pct',
        y.var='pct',
        fill.var='is.loop.shared',
        label.var='pct.label',
        facet.row='comparison',
        position='stack',
        # scales='fixed',
        # legend.position='right',
        strip.text=element_text(size=10),
        axis.text.x=element_text(angle=45, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()
    )
```

# Compare Loop Stats by Reproducibility

## Genome-Wide {.tabset}

```{r idr2d_compare_stats_gw, results='asis', fig.dim=c(10,6), eval=FALSE}
idr2d.results.df %>%
    make_nested_plot_tabs(
    )
```

## Per Chr {.tabset}

```{r idr2d_compare_stats_chr, results='asis', fig.dim=c(12,10), eval=FALSE}
idr2d.results.df %>%
    make_nested_plot_tabs(
    )
```

