---
title: "16p HiC - TADCompare Analysis"
author: "Siddharth Reed"
date: "`r Sys.Date()`"
---
<style type="text/css">
.main-container {
  max-width: 1800px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Dependencies + Load Data

```{r knitr}
library(knitr)
knitr::opts_chunk$set( 
    results=FALSE,
    message=FALSE,
    warning=FALSE,
    dev=c('png', 'pdf', 'svg'),
    dpi=300
)
```
Set up file locations + load dependences
```{r dependencies}
library(here)
BASE_DIR <- here()
here::i_am('notebooks/TADCompare.Rmd')
source(file.path(BASE_DIR, 'scripts', 'locations.R'))
source(file.path(SCRIPT_DIR, 'utils.data.R'))
source(file.path(BASE_DIR, 'scripts', 'constants.R'))
source(file.path(SCRIPT_DIR, 'utils.plot.R'))
source(file.path(SCRIPT_DIR, 'TADs', 'utils.TADs.R'))
source(file.path(SCRIPT_DIR, 'TADs', 'utils.TADCompare.R'))
library(tidyverse)
library(magrittr)
```
Load TADCompare results for boundaries only
```{r load_TADCompare_data}
TADCompare.df <- 
    check_cached_results(
        results_file=TADCOMPARE_RESULTS_FILE,
        # force_redo=TRUE,
        results_fnc=load_all_TADCompare_results,
        gw.fdr.threshold=1,
        fdr.threshold=1,
        nom.threshold=0.05
    ) %>%
    post_process_TADCompare_results() %>% 
    mutate(log.p.adj.gw=-log10(p.adj.gw)) %>% 
    select(-c(TAD.isDifferential, isDifferential)) %>% 
    filter(isBoundary == 'TAD') %>% 
    standardize_data_cols()
```
Load summarized results
```{r load_TADCompare_freq_data}
n.TADCompare.df <- 
    check_cached_results(
        results_file=TADCOMPARE_COUNTS_RESULTS_FILE,
        # force_redo=TRUE,
        results_fnc=load_correct_count_all_TADCompare_results
    ) %>%
    post_process_TADCompare_results() %>% 
    filter(!is.na(DifferenceType)) %>% 
    filter(isBoundary == 'TAD') %>% 
    standardize_data_cols()
```

# Plot TAD differences

## All bins Considered {.tabset}

### Genome-Wide {.tabset}

```{r n_Differences_heatmap_gw, results='asis', fig.dim=c(7,3)}
# Plot pct barchart of diff vs non-diff per sig lvl
n.TADCompare.df %>% 
    calc_pct(
        c(
            'chr',
            'DifferenceType',
            'Enriched.Condition'
        )
    ) %>% 
    make_nested_plot_tabs(
        group.cols=c('resolution'),
        max.header.lvl=4,
        plot.fnc=plot_heatmap,
        y.var='comparison',
        x.var='sig.lvl',
        fill.var='n',
        fill.scale.mode='m',
        label.var='n',
        label.size=2,
        facet.col='TAD.method',
        facet.row='isBoundary',
        scales='fixed',
        # legend.position='right',
        strip.text=element_text(size=8),
        axis.text.x=element_text(angle=35, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()
    )
```

### Per Chr{.tabset}

```{r n_Differences_heatmap_chr, results='asis', fig.dim=c(10,6)}
n.TADCompare.df %>% 
    calc_pct(
        c(
            'DifferenceType',
            'Enriched.Condition'
        )
    ) %>% 
    make_nested_plot_tabs(
        # group.cols=c('resolution'),
        group.cols=c('resolution', 'TAD.method'),
        max.header.lvl=4,
        plot.fnc=plot_heatmap,
        x.var='chr',
        y.var='sig.lvl',
        fill.var='n',
        fill.scale.mode='m',
        label.var='n',
        label.size=2,
        facet.row='comparison',
        scales='fixed',
        # legend.position='right',
        strip.text=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()
    )
```

## Differential Boundaries Only {.tabset}

### Genome-Wide {.tabset}

```{r n_Differences_barplot_gw, results='asis', fig.dim=c(10,6)}
n.TADCompare.df %>% 
    filter(sig.lvl != 'N.S.') %>% 
    calc_pct(
        c(
            'chr',
            'sig.lvl'
            # 'Enriched.Condition'
        )
    ) %>% 
    make_nested_plot_tabs(
        group.cols=c('resolution'),
        max.header.lvl=4,
        plot.fnc=plot_barplot,
        x.var='Enriched.Condition',
        y.var='pct',
        y.scale.mode='pct',
        position='stack',
        fill.var='DifferenceType',
        label.var='n.and.pct.label',
        label.size=2,
        facet.col='comparison',
        facet.row='TAD.method',
        scales='free_x',
        strip.text=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1),
        # axis.title.x=element_blank(),
        axis.title.y=element_blank()
    )
```

### Per Chr{.tabset}

```{r n_Differences_barplot_chr, results='asis', fig.dim=c(10,6)}
n.TADCompare.df %>% 
    filter(sig.lvl != 'N.S.') %>% 
    calc_pct(
        c(
            # 'chr',
            'sig.lvl',
            'Enriched.Condition'
        )
    ) %>% 
    make_nested_plot_tabs(
        group.cols=c('resolution', 'TAD.method'),
        max.header.lvl=4,
        plot.fnc=plot_barplot,
        x.var='chr',
        y.var='pct',
        y.scale.mode='pct',
        facet.row='comparison',
        # y.var='pct',
        # y.scale.mode='pct',
        # x.var='comparison',
        # facet.group='chr',
        # facet.nrow=4,
        position='stack',
        fill.var='DifferenceType',
        label.var='n.and.pct.label',
        label.size=2,
        # scales='free_x',
        strip.text=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()
    )
```

# Significance Box Plots {.tabset}

```{r n_Differences_boxplot_gw, results='asis', fig.dim=c(10,6)}
TADCompare.df %>% 
    make_nested_plot_tabs(
        group.cols=c('resolution'),
        max.header.lvl=2,
        plot.fnc=plot_boxplot,
        x.var='DifferenceType',
        y.var='log.p.adj.gw',
        fill.var='comparison',
        facet.col='TAD.method',
        # scales='free_x',
        strip.text=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1),
        axis.title.x=element_blank()
        # axis.title.y=element_blank()
    )
```

# Significance by type {.tabset}

```{r n_Differences_boxplot_by_type_gw, results='asis', fig.dim=c(10,6)}
TADCompare.df %>% 
    make_nested_plot_tabs(
        group.cols=c('resolution'),
        max.header.lvl=2,
        plot.fnc=plot_boxplot,
        x.var='comparison',
        y.var='log.p.adj.gw',
        fill.var='DifferenceType',
        facet.row='TAD.method',
        # scales='free_x',
        strip.text=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1),
        axis.title.x=element_blank()
        # axis.title.y=element_blank()
    )
```

# TAD Difference Positions {.tabset}

```{r position_scatter_chr, results='asis', fig.dim=c(10,6), eval=FALSE}
TADCompare.df %>% 
    select(
        resolution,
        TAD.params,
        TAD.method, 
        comparison,
        chr,
        Boundary,
        isBoundary, DifferenceType,
        TAD.Score.Numerator, TAD.Score.Denominator, 
        Gap.Score,
        # p.value, p.adj
        # p.adj.gw,
        log.p.adj.gw
    ) %>% 
    # gap score lineplot + point with difference type
    make_nested_plot_tabs(
        group.cols=c('resolution', 'TAD.method',  ' chr'),
        max.header.lvl=3,
        plot.fnc=plot_lineplot,
        x.var='Boundary',
        x.scale.mode='mb',
        y.var='Gap.Score',
        color.var='comparison',
        # scales='free_x',
        strip.text=element_text(size=8),
        axis.text.x=element_text(angle=45, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank()
    )
```

