---
title: "TAD Analysis for 16p"
author: "Siddharth Reed"
date: "`r Sys.Date()`"
output: 
    html_document:
        keep_md: yes
        toc: true
        toc_float: true
        theme: paper
        code_folding: hide
        df_print: paged
---
<style type="text/css">
.main-container {
  max-width: 1800px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Set Up

```{r knitr}
library(knitr)
knitr::opts_chunk$set(
    dev=c('png', 'pdf', 'svg'),
    dpi=300
)
```
Set up file locations + load dependences
```{r dependencies, results=FALSE, message=FALSE, warning=FALSE}
library(here)
here::i_am('notebooks/TADs.Rmd')
BASE_DIR <- here()
source(file.path(here(), 'scripts', 'locations.R'))
source(file.path(SCRIPT_DIR, 'constants.R'))
source(file.path(SCRIPT_DIR, 'utils.data.R'))
source(file.path(SCRIPT_DIR, 'utils.plot.R'))
source(file.path(SCRIPT_DIR, 'utils.TADs.R'))
library(tidyverse)
library(magrittr)
library(ggplot2)
library(ggh4x)
library(purrr)
```

# hiTAD TAD annotations {.tabset}

Load all hiTAD annotations
```{r load_hiTAD_TADs}
sample.metadata <- load_sample_metadata()
all.hitad.annotations <- 
    check_cached_results(
        results_file=TAD_RESULTS_FILE,
        results_fnc=load_all_TAD_annotations,
        # force_redo=TRUE,
        pattern='/method_hiTAD/'
    ) %>% 
    inner_join(
        sample.metadata %>% select(SampleID),
        by=join_by(SampleID)
    ) %>% 
    mutate(
        isMerged=factor(ifelse(isMerged, 'Merged', 'Individual')),
        resolution=scale_numbers(resolution),
        chr=factor(chr, levels=CHROMOSOMES),
        TAD.length=TAD.end - TAD.start
    ) %>% 
    select(
        isMerged,
        Edit,
        Genotype,
        Celltype,
        CloneID,
        TechRepID,
        SampleID,
        resolution,
        weight,
        chr,
        TAD.start,
        TAD.end,
        TAD.length
    )
# Subset to only relevant parameters
hitad.annotations <- 
    all.hitad.annotations %>% 
    filter(weight == 'ICE')
```

## Number of TADs

```{r PLOT_hitad_tad_number, results='asis', fig.width=8, fig.height=6}
all.hitad.annotations %>% 
    dplyr::count(
        isMerged, weight, Celltype, 
        resolution, Genotype,
        SampleID, chr, name='nTADs'
    ) %>% 
    make_nested_plot_tabs(
        group_cols=c('isMerged', 'weight', 'resolution'),
        # group_cols=c('isMerged', 'resolution', 'weight'),
        max_header_lvl=3,
        plot_fnc=plot_heatmap,
        x_var='chr',
        y_var='SampleID',
        fill_var='nTADs',
        # facet_col='Genotype',
        facet_row='Celltype',
        scales='free_y',
        axis.text.x=element_text(angle=45, hjust=1)
    )
```

## TAD Lengths

```{r PLOT_hitad_tad_length, results='asis', fig.width=8, fig.height=6}
all.hitad.annotations %>% 
    make_nested_plot_tabs(
        group_cols=c('isMerged', 'weight', 'resolution'),
        max_header_lvl=3,
        plot_fnc=plot_boxplot,
        x_var='chr',
        y_var='TAD.length',
        fill_var='Genotype',
        facet_row='Celltype',
        scales='fixed',
        scale_mode='mb',
        axis.text.x=element_text(angle=45, hjust=1)
    )
```

## TAD Similarity

Calculate MoC for all pairs of TAD annotation sets
```{r hiTAD_calculate_mocs}
hitad.mocs <- 
    check_cached_results(
        results_file=TAD_MOC_FILE,
        # force_redo=TRUE,
        results_fnc=calculate_all_pairs_MoCs,
        tad.annotations=
            hitad.annotations %>%
            nest(
                TADs=
                    c(
                        TAD.start,
                        TAD.end,
                        TAD.length
                    )
            ) %>% 
            nest(
                SampleInfo=
                    c(
                        weight,
                        SampleID,
                        Edit,
                        Genotype,
                        Celltype,
                        CloneID,
                        TechRepID
                    )
            ),
        pair_grouping_cols=
            c(
                'isMerged',
                'resolution',
                'chr'
            )
    ) %>% 
    # mutate(resolution=scale_numbers(resolution)) %>% 
    mutate(chr=factor(chr, levels=CHROMOSOMES)) %>% 
    rename('Measure.of.Concordance'=mocs)
# hitad.mocs
```
Plot distribution of similarity between genotype pairs 
```{r plot_MoCs_boxplot, results='asis'}
hitad.mocs %>% 
    make_nested_plot_tabs(
        group_cols=c('isMerged', 'weight', 'resolution'),
        max_header_lvl=3,
        plot_fnc=plot_boxplot,
        x_var='Genotype',
        y_var='Measure.of.Concordance',
        facet_col='Celltype',
        # facet_row='resolution',
        scales='fixed',
        axis_label_accuracy=1,
        axis.text.x=element_text(angle=45, hjust=1)
    )
```
Plot mean similarity between genotype pairs per chromosome
```{r plot_MoCs_heatmap, results='asis'}
hitad.mocs %>% 
    group_by(weight, resolution, isMerged, chr, Genotype, Celltype) %>%
    summarize(mean.MoC=mean(Measure.of.Concordance)) %>% 
    make_nested_plot_tabs(
        group_cols=c('isMerged', 'weight', 'resolution'),
        max_header_lvl=3,
        plot_fnc=plot_heatmap,
        x_var='chr',
        y_var='Genotype',
        fill_var='mean.MoC',
        facet_row='Celltype',
        # facet_row='resolution',
        scales='fixed',
        legend.position='top',
        axis.text.x=element_text(angle=45, hjust=1)
    )
```

# cooltools Boundary Annotations

## Load cooltools annotations

- Calling Parameters
  - resolution
  - weight
  - threshold
  - mfvp
  - window.size
- Individual TAD data
  - chr
  - TAD.start
  - TAD.end
  - n_valid_pixels
  - boundary_strength
```{r load_cooltools_TADs, eval=FALSE}
# tad.annotations %>% colnames()
# tad.annotations %>% filter(region != chr)
# Cooltools annotations
all.cooltools.annotations <- 
    tad.annotations %>%
    filter(method == 'cooltools') %>% 
    rename('TAD.boundary'=TAD.start) %>% 
    mutate(
        window.size.bins=window.size / resolution,
        window.size=scale_numbers(window.size),
        resolution=scale_numbers(resolution)
    ) %>% 
    select(-c(TAD.end, window.size, method))
    # filter(ReadFilter == 'mapq_30') %>% 
    # select(-c(ReadFilter, region, is_boundary)) %>% 
    get_info_from_SampleIDs(
        sample_ID_col='SampleID',
        keep_id=TRUE,
        nest_col=NA
    )
# Filter tads to relevant parameter set
cooltools.annotations <- 
    all.cooltools.annotations %>%
    filter(
        weight == 'ICE',
        mfvp == 0.9,
        window.size.bins == 60,
        threshold == 'Li'
    )
```

## cooltools parameter variability

Now plot number of TADs/boundaries 
```{r PLOT_cooltools_tad_number, results='asis', fig.width=8, fig.height=6, eval=FALSE}
cooltools.annotations %>% 
    mutate(resolution=scale_numbers(resolution)) %>% 
    group_by(isMerged, resolution, Sample.ID, chr) %>% 
    dplyr::count(name='nBoundaries') %>% 
    make_nested_plot_tabs(
        group_cols=c('isMerged'),
        max_header_lvl=3,
        plot_fnc=plot_heatmap,
        x_var='chr',
        y_var='Sample.ID',
        fill_var='nBoundaries',
        facet_row='resolution',
        scales='fixed'
    )
    # group_by(isMerged) %>%
    # group_map(
    #     ~ plot_heatmap(
    #         plot.df=.x,
    #         x_var='chr',
    #         y_var='Sample.ID',
    #         fill_var='nBoundaries',
    #         facet_row='resolution',
    #         scales='fixed'
    #     )
    # ) %>% 
    # cowplot::plot_grid(
    #     plotlist=.,
    #     align='v',
    #     ncol=1
    # ) %>% 
    # ggsave(
    #     file.path(PLOT_DIR, 'cooltools.nBoundariess.heatmap.pdf'),
    #     .,
    #     width=10,
    #     height=10,
    #     units='in'
    # )
```
Number of boundaries per chr + sample across param combos
```{r PLOT_boundaries_per_paramset, results='asis', fig.width=8, fig.height=6, eval=FALSE}
all.cooltools.annotations %>% 
    mutate(resolution=scale_numbers(resolution)) %>% 
    mutate(params=glue('{weight},{mfvp},{window.size.bins}')) %>% 
    group_by(
        across(
            -c(
               chr,
               TAD.boundary,
               n_valid_pixels,
               boundary_strength
            )
        )
    ) %>% 
    dplyr::count(name='nBoundaries') %>%
    ungroup() %>% 
    make_nested_plot_tabs(
        group_cols=c('isMerged'),
        max_header_lvl=3,
        plot_fnc=plot_boxplot,
        x_var='params',
        y_var='nBoundaries',
        fill_var='isMerged',
        facet_col='threshold',
        facet_row='resolution',
        scales='free_y',
    )
    # plot_boxplot(
    #     x_var='params',
    #     y_var='nBoundaries',
    #     fill_var='isMerged',
    #     facet_col='threshold',
    #     facet_row='resolution',
    #     scales='free_y'
    # ) %>%
    # ggsave(
    #     file.path(PLOT_DIR, 'cooltools.boundaries_per_paramset.pdf'),
    #     .,
    #     width=10,
    #     height=10,
    #     units='in'
    # )
```
Strength of annotated boundaries per across param comvos
```{r PLOT_boundary_strength_per_paramset, results='asis', fig.width=8, fig.height=6, eval=FALSE}
all.cooltools.annotations %>% 
    mutate(resolution=scale_numbers(resolution)) %>% 
    mutate(params=glue('{weight},{mfvp},{window.size.bins}')) %>% 
    make_nested_plot_tabs(
        group_cols=c('isMerged'),
        max_header_lvl=3,
        plot_fnc=plot_boxplot,
        x_var='params',
        y_var='boundary_strength',
        fill_var='isMerged',
        facet_col='threshold',
        facet_row='resolution',
        scales='fixed'
    )
    # plot_boxplot(
    #     x_var='params',
    #     y_var='boundary_strength',
    #     fill_var='isMerged',
    #     facet_col='threshold',
    #     facet_row='resolution',
    #     scales='fixed'
    # ) %>%
    # ggsave(
    #     file.path(PLOT_DIR, 'cooltools.boundary_strength_per_paramset.pdf'),
    #     .,
    #     width=10,
    #     height=10,
    #     units='in'
    # )
```
Choose relevant parameter sets to use

# Load Directionality Index scores

```{r load_DI, eval=FALSE}
DI.scores <- 
    load_all_TAD_annotations(
        tad_annotations_dir=glue('{BASE_DIR}/results/TADs/hiTAD'),
        tad.method='hiTAD-DIs',
        file_suffix='-DIs.txt',
        param_names=
            c(
                'Normalization',
                'Resolution',
                'SampleName'
            ),
    ) %>%
    separate_wider_delim(
        SampleName,
        delim=fixed('.'),
        names=
            c(
                'SampleName',
                NA,
                'ReadFilter',
                NA
            )
    ) %>% 
    mutate(
        Resolution=as.integer(Resolution),
        chr=factor(chr, levels=CHROMOSOMES)
    ) 
DI.scores %>% head(2) %>% t()
```

# Plot DI Scores

```{r plot_DI, eval=FALSE}
tmp <- 
    DI.scores %>%
    add_count(Resolution, SampleName, chr, name='n.bins') %>% 
    left_join(
        tad.annotations %>% select(c(Resolution, SampleName, chr, TAD.start, TAD.end)),
        by=
            join_by(
                Resolution,
                chr,
                SampleName,
                between(bin.start, TAD.start, TAD.end, bounds='[)')
            ) 
    ) %>%
    mutate(
        TAD.boundaries=
            case_when(
                bin.start == TAD.start ~ 'TAD.Start',
                bin.end == TAD.end ~ 'TAD.End',
                bin.start >= TAD.start & bin.end <= TAD.end ~ 'TAD.inner',
                TRUE ~ 'Non-TAD'
        )
    )
tmp %>% 
    # filter(chr %in% paste0('chr', c(1,3,5,10,16,20,'X'))) %>% 
    filter(chr == 'chr16') %>% 
    mutate(Resolution=glue('{Resolution / 1000}Kb')) %>%
    mutate(Resolution=factor(Resolution, levels=c('100Kb', '50Kb', '10Kb', '5Kb'))) %>% 
    group_by(Method) %>%
    group_map(
        ~ ggplot(
            .x, 
            aes(
                x=bin.start,
                y=DI,
                group=chr,
                color=Genotype
                # color=Genotype, shape=Replicate
            )
        ) +
        # geom_rect(data=.x %>% filter(SampleName == ''), aes(xmin=TAD.start, xmax=TAD.end, ymin=-Inf, ymax=0, fill=SampleName), alpha=0.01, size=0) +
        geom_line() +
        labs(x='bins', y='Directionality Index') +
        ggh4x::facet_grid2(cols=vars(chr), rows=vars(Resolution), scales='free_y') +
        scale_x_continuous(
            n.breaks=10,
            expand=c(0.01, 0.01, 0.01, 0.01),
            limits=c(0, NA)
        ) +
        theme(
            legend.position='top',
            axis.title.x=element_blank(),
            axis.text.x=element_text(angle=45, hjust=1),
            axis.text.y=element_markdown()
        ) +
        add_ggtheme(),
        .keep=TRUE
    ) %>% 
    cowplot::plot_grid(
        plotlist=.,
        align='v',
        ncol=1
    )
```

# Scratch

```{r print_stuff, eval=FALSE}
cooltools.annotations
cooltools.annotations %>% colnames()
hitad.annotations %>% colnames()
hitad.annotations
hitad.mocs
```

